---
title: "p8105_hw1_are2132"
author: "Alison Elgass"
date: 2019-09-15
output: github_document
---

# Problem 1
First we create a dataframe `df1` with 4 vectors, 
each of size 8  
Each vector utilizes a different variable type

```{r createdf1}
library(tidyverse)
df1 = tibble(
  normSamp = rnorm(8),                      #normal sample
  logicVect1 = normSamp > 0,                #logical
  charVect1 = character(length = 8),        #character
  factorVect1 = factor(c(1,2,3,1,2,3,1,2))  #factor
)
```

## Changing types
Next we use the as.numeric() function to convert vectors
```{r convert, eval=FALSE}
a = as.numeric(pull(df1,logicVect1))
b = as.numeric(pull(df1,charVect1))
c = as.numeric(pull(df1,factorVect1))
```
This exercise shows that the logical and factor vectors can be converted into numerical vectors (a and c) from which we can take their mean, but the character vector cannot (vector b shows just 8 'NA's)  

## Some multiplication
Now we convert the logical vector 3 ways, then multiply by our random sample
```{r prob1math}
as.numeric(pull(df1,logicVect1))*pull(df1,normSamp)
as.factor(pull(df1,logicVect1))*pull(df1,normSamp)
as.numeric(as.factor(pull(df1,logicVect1)))*pull(df1,normSamp)
```
Again, we see that in the middle example above we _cannot_ 
get a reasonable numerical output since we are attempting 
to multiply a numeric by a factor  


# Problem 2
First we create a dataframe `df2` with 5 vectors
```{r createdf2}
df2 = tibble(
  x = rnorm(500), #sample 500 from normal
  y = rnorm(500),
  logicVect2 = (x + y) > 1,
  numerVect2 = as.numeric(logicVect2),
  factorVect2 = as.factor(logicVect2)
)
```
In summary,  
The dataset `df2` has `r nrow(df2)` rows and `r ncol(df2)` columns  
The mean of `x` is `r mean(pull(df2,x))`  
The median of `x` is `r median(pull(df2,x))`  
The standard deviation of `x` is `r sd(pull(df2,x))`  
The proportion of cases for which `x + y > 1` is
`r sum(pull(df2,logicVect2))/nrow(df2)`

## Scatterplot 1

```{r scatterplot1}
ggplot(df2, aes(x = x, y = y, color = logicVect2)) + geom_point()
```
    
This scatterplot shows y vs. x from our dataframe `df2`  
The color of each point corresponds to the value of the logical vector for the (x,y) pair. The point is  
  
*  Blue if `(x+y) > 1` or `logicVect2 = TRUE`  
*  Red if `(x+y) <= 1` or `logicVect2 = FALSE`  
  
  
We'll also **save this plot** to the R project as a 
jpeg file
```{r saveplot1}
plot1 = ggplot(df2, aes(x = x, y = y, color = numerVect2)) + geom_point()
ggsave("scatter_plot.jpeg", plot1)
```
  
## Scatterplot 2
```{r scatterplot2}
ggplot(df2, aes(x = x, y = y, color = numerVect2)) + geom_point()
```
  
In this plot, each point is  
  
*  Blue if `(x+y) > 1` or `numerVect2 = 1`  
*  Red if `(x+y) <= 1` or `numerVect2 = 0`  
  
## Scatterplot 3
```{r scatterplot3}
ggplot(df2, aes(x = x, y = y, color = factorVect2)) + geom_point()
```
  
In this plot, each point is  
  
*  Blue if `(x+y) > 1` or `factorVect2 = TRUE`  
*  Red if `(x+y) <= 1` or `factorVect2 = FALSE` 
  


