---
title: "p8105_hw1_are2132"
author: "Alison Elgass"
output: github_document
---

# Problem 1
First we create a dataframe with 4 vectors

```{r createdf1}
library(tidyverse)
df1 = tibble(
  normSamp = rnorm(8),                     #normal sample
  logicVect1 = normSamp > 0,               #logical
  charVect1 = character(length = 8),       #character
  factorVect1 = factor(c(1,2,3,1,2,3,1,2)) #factor
)
```

Use the as.numeric function to convert vectors
```{r convert, eval=FALSE}
a = as.numeric(pull(df1,logicVect1))
b = as.numeric(pull(df1,charVect1))
c = as.numeric(pull(df1,factorVect1))
```

Convert logical vector 3 ways, then multiply by random sample
```{r prob1math}
as.numeric(pull(df1,logicVect1))*pull(df1,normSamp)
as.factor(pull(df1,logicVect1))*pull(df1,normSamp)
as.numeric(as.factor(pull(df1,logicVect1)))*pull(df1,normSamp)
```

# Problem 2
First we create a dataframe with 5 vectors
```{r createdf2}
df2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logicVect2 = (x + y) > 1,
  numerVect2 = as.numeric(logicVect2),
  factorVect2 = as.factor(logicVect2)
)
```

The dataset `df2` has `r nrow(df2)` rows and `r ncol(df2)` columns  
The mean of `x` is `r mean(pull(df2,x))`  
The median of `x` is `r median(pull(df2,x))`  
The standard deviation of `x` is `r sd(pull(df2,x))`  
The proportion of cases for which x + y > 1 is
`r sum(pull(df2,logicVect2))/nrow(df2)`

```{r scatterplot}
ggplot(df2, aes(x = x,y = y, color=logicVect2)) + geom_point()
```

